---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
The intention of this study is to process and analyze the information related to 
the activities being done by a subject in periods of time. This is now possible with the 
rise of new technologies and devices that let us access data that was too complex to get obtain in the past. In particular this study works in the analysis of 5 minute intervals that monitor the amount of steps of an anonymous individual each day from October to November 2012.

## Loading and preprocessing the data
The data is available is the activity.zip file. and can be loaded with the following instructions
```{r loadProcess}
unzip("activity.zip")
activities <- read.csv("activity.csv")
```
Many of the questions answered here refer to daily information, the information was grouped 
and processed in a daily basis so we group and sum the amount of steps by day using the dplyr library.

```{r message=FALSE}
library(dplyr)
daily <- group_by(activities,date,steps)
daily <- summarize(daily)
```

## What is mean total number of steps taken per day?

The distribution of steps per day along the duration of the experiment is described by the following histogram

```{r}
hist(daily$steps,
     xlab = "Total Steps By day",
     col="blue")
```


After arranging the number of steps by day we can obtain the means of the steps the subject executed during the experiment


```{r}
  mean(daily$steps,na.rm = T)
  median(daily$steps,na.rm =T)
```




## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
